Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-02-23 analysis/paper_draft3.Rmd
revised TAS memo - more on hyperprior for alpha etc.; updated references, minor edit to paper
edit to discussion

In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, 2243; Wash. General Rule 37(g)(v); Calif. Code Civil Procedure $\S$ 231.7(d)(3)(G)]. Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex. Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013]. Prior studies have also modeled how much a prospective juror's race affected the odds of being struck. For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2022.

@@

In general, by their choice of $\alpha$, the lawyer explicitly and transparently weights the historical strike data in the analysis. This transparency thwarts lawyer efforts to strategically select $\alpha$ values. Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of $\alpha$. In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation). Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g. @Ibrahim2015, 3724-3738] could appropriately apply in this context. Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g. @Han2023; @Pawel2022].

>>>>

In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, p. 2243; Wash. General Rule 37(g)(v); Calif. Code Civil Procedure $\S$ 231.7(d)(3)(G)]. Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex. Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013]. Prior studies have also modeled how much a prospective juror's race affected the odds of being struck. For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2022.

@@

In general, by their choice of $\alpha$, the lawyer explicitly and transparently weights the historical strike data in the analysis. This transparency thwarts lawyer efforts to strategically select $\alpha$ values. Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of $\alpha$. In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation). Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g. @Ibrahim2015, 3724-3738] could appropriately apply in this context. while also considering incompatibility (commeasureability) in selecting $\alpha$ (@Ollier2020). Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g. @Han2023; @Pawel2022].
